---
title: "infection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{infection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(culex)
library(deSolve)
library(parallel)
library(data.table)
library(ggplot2)
```

We'll show how to use the mosquito model with infection here. The mosquito model
with infection is designed to be integrated into a larger model framework, where the 
different components pass information relevant to pathogen transmission every time
step.

A good way to check that our infection model is working correctly is to check that
in the absence of infection, it returns the same results as the lifecycle model.

We will use the same parameters as in the vignette "Example", please see that file for
explanation of parameters and simulation workflow.


```{r}
parameters <- culex_parameters()

dt <- 0.1
tmax <- 365*3
maxstep <- tmax/dt

# initial # of adults
A0 <- c(600, 0)

# movement between 2 patches
psi <- matrix(
  data = c(0.95, 0.05,
           0.025, 0.975),
  nrow = 2, ncol = 2, byrow = TRUE
)

# delays at t=0
tau0 <- c("E"=0,"L"=0,"P"=0,"EIP"=0)
temp0 <- temperature(0, parameters)
tau0[1] = 1 / egg_maturation_rate(temp0, parameters) # tau_E
tau0[2] = 1 / larvae_maturation_rate(temp0, parameters) # tau_L
tau0[3] = 1 / pupae_maturation_rate(temp0, parameters) # tau_P
tau0[4] = 1 / eip_rate(temp = temp0, parameters)

# integrate past the simulation end time by a comfortable amount
times <- seq(from=0.0,to=tmax+200,by=dt)
tau_ode <- deSolve::ode(y = tau0, times = times, func = tau_diffeqn, parms = parameters, method = "ode23")

tau_traj <- as.data.table(tau_ode)
tau_traj[, "step" := seq_along(times)]
tau_traj[, c("E", "L", "P", "EIP") := lapply(.SD, function(x){as.integer(round(x = x/dt))}), .SDcols = c("E", "L", "P", "EIP")]
tau_traj[, "time" := NULL]

# go from backward-looking maturation delays to forward-looking delays
tau_traj[, "step_forward_E" := as.integer(step - E)]
tau_traj[, "step_forward_L" := as.integer(step - L)]
tau_traj[, "step_forward_P" := as.integer(step - P)]
tau_traj[, "step_forward_EIP" := as.integer(step - EIP)]

tauE <- as.integer(tau_traj[step_forward_E %in% 1:maxstep, "E"][[1]])
tauL <- as.integer(tau_traj[step_forward_L %in% 1:maxstep, "L"][[1]])
tauP <- as.integer(tau_traj[step_forward_P %in% 1:maxstep, "P"][[1]])
tauEIP <- as.integer(tau_traj[step_forward_EIP %in% 1:maxstep, "EIP"][[1]])
```

Let's draw the deterministic trajectories. Because we don't set `f`, `q`, or `kappa`,
they will all be zero and no infection should occur.

```{r}
# solve a determiinistic trajectory
mod <- create_culex_deterministic(p = 2, tau_E = tauE, tau_L = tauL, tau_P = tauP, dt = dt, psi = psi)
set_A_deterministic(mod = mod, A = A0)

out_d <- data.table(day = rep(1:tmax, each = 8), stage = rep(c("E","L","P","A"), tmax*2), patch = rep(c(1,1,1,1,2,2,2,2), tmax), value = NaN)
setkey(out_d, "day")
out_d[, "patch" := as.factor(patch)]
out_i <- 1L

for (i in 1:maxstep) {
  step_culex_deterministic(mod = mod, parameters = parameters)
  if ((i-1) %% (1/dt) == 0) {
    out_d[day == out_i & stage == "E", "value" := as.vector(get_E_deterministic(mod))]
    out_d[day == out_i & stage == "L", "value" := as.vector(get_L_deterministic(mod))]
    out_d[day == out_i & stage == "P", "value" := as.vector(get_P_deterministic(mod))]
    out_d[day == out_i & stage == "A", "value" := as.vector(get_A_deterministic(mod))]
    out_i <- out_i + 1L
  }
}

# solve a deterministic trajectory from model with infection dynamics
mod <- create_culex_infection_deterministic(p = 2, tau_E = tauE, tau_L = tauL, tau_P = tauP, tau_EIP = tauEIP, dt = dt, psi = psi, n_species = 1)
set_AS_infection_deterministic(mod = mod, A = A0)

out_d_inf <- data.table(day = rep(1:tmax, each = 8), stage = rep(c("E","L","P","A"), tmax*2), patch = rep(c(1,1,1,1,2,2,2,2), tmax), value = NaN)
setkey(out_d_inf, "day")
out_d_inf[, "patch" := as.factor(patch)]
out_i <- 1L

for (i in 1:maxstep) {
  step_culex_infection_deterministic(mod = mod, parameters = parameters)
  if ((i-1) %% (1/dt) == 0) {
    out_d_inf[day == out_i & stage == "E", "value" := colSums(get_E_infection_deterministic(mod))]
    out_d_inf[day == out_i & stage == "L", "value" := colSums(get_L_infection_deterministic(mod))]
    out_d_inf[day == out_i & stage == "P", "value" := colSums(get_P_infection_deterministic(mod))]
    out_d_inf[day == out_i & stage == "A", "value" := as.vector(get_AS_infection_deterministic(mod))]
    out_i <- out_i + 1L
  }
}

out_d[, "Model" := "Lifecycle"]
out_d_inf[, "Model" := "Infection"]
out_d <- rbind(out_d, out_d_inf)

ggplot(out_d) +
  geom_line(aes(x=day,y=value,color=Model,linetype=patch), alpha = 0.5, size = 1.25) +
  facet_wrap(. ~ stage, scales = "free") +
  theme_bw()
```

We will draw 10 samples from the stochastic model.

```{r}
out <- parallel::mclapply(X = 1:10, FUN = function(runid) {
  
  # lifecycle
  mod <- create_culex_stochastic(p = 2, tau_E = tauE, tau_L = tauL, tau_P = tauP, dt = dt, psi = psi)
  set_A_stochastic(mod = mod, A = A0)

  out <- data.table(day = rep(1:tmax, each = 8), stage = rep(c("E","L","P","A"), tmax*2), patch = rep(c(1,1,1,1,2,2,2,2), tmax), value = NaN)
  setkey(out, "day")
  out[, "patch" := as.factor(patch)]
  out_i <- 1L

  for (i in 1:maxstep) {
    step_culex_stochastic(mod = mod, parameters = parameters)
    if ((i-1) %% (1/dt) == 0) {
      out[day == out_i & stage == "E", "value" := as.vector(get_E_stochastic(mod))]
      out[day == out_i & stage == "L", "value" := as.vector(get_L_stochastic(mod))]
      out[day == out_i & stage == "P", "value" := as.vector(get_P_stochastic(mod))]
      out[day == out_i & stage == "A", "value" := as.vector(get_A_stochastic(mod))]
      out_i <- out_i + 1L
    }
  }
  
  # infection
  mod <- create_culex_infection_stochastic(p = 2, tau_E = tauE, tau_L = tauL, tau_P = tauP, tau_EIP = tauEIP, dt = dt, psi = psi, n_species = 1)
  set_AS_infection_stochastic(mod = mod, A = A0)
  
  out_inf <- data.table(day = rep(1:tmax, each = 8), stage = rep(c("E","L","P","A"), tmax*2), patch = rep(c(1,1,1,1,2,2,2,2), tmax), value = NaN)
  setkey(out_inf, "day")
  out_inf[, "patch" := as.factor(patch)]
  out_i <- 1L
  
  for (i in 1:maxstep) {
    step_culex_infection_stochastic(mod = mod, parameters = parameters)
    if ((i-1) %% (1/dt) == 0) {
      out_inf[day == out_i & stage == "E", "value" := colSums(get_E_infection_stochastic(mod))]
      out_inf[day == out_i & stage == "L", "value" := colSums(get_L_infection_stochastic(mod))]
      out_inf[day == out_i & stage == "P", "value" := colSums(get_P_infection_stochastic(mod))]
      out_inf[day == out_i & stage == "A", "value" := as.vector(get_AS_infection_stochastic(mod))]
      out_i <- out_i + 1L
    }
  }
  
  out[, "model" := "Lifecycle"]
  out_inf[, "model" := "Infection"]
  
  out <- rbind(out, out_inf)
  out[ , "run" := as.integer(runid)]
  
  return(out)
})


```

Plot the simulations.

```{r}
out_sum <- do.call(rbind, out)

ggplot(out_sum) +
  geom_line(aes(x=day,y=value,color=model,linetype=patch), alpha = 0.35) +
  facet_wrap(. ~ stage, scales = "free") +
  theme_bw()
```
